---
title: "Group by (salaries)"
output:
  html_document:
    df_print: paged
---

```{r}
library(tidyverse)
library(readxl)
```

Load data
```{r}
salaries <- read_excel("data/salaries.xlsx", sheet=1)
```

Run integrity checks on columns in your data

Use the `count()` function to look for misspellings and inconsistencies (and also to see which values are most common)
```{r}
salaries %>% count(department)
```

These two lines of code produce exactly the same results.
```{r}
salaries %>% count(job_title)
salaries %>% group_by(job_title) %>% summarise(n = n())
```

Check the minimum and maximum of numbers columns like total_comp:
```{r}
salaries %>% summarise(min(total_comp), max(total_comp))
```

Which employee made the most amount of money (hint: total compensation)? 
```{r}
salaries %>% arrange(desc(total_comp))
```

What is the total compensation paid out by the city in 2021? 
```{r}
salaries %>% summarise(overtime = sum(overtime_oncall, na.rm=TRUE))
```

Which department paid out the most in overtime pay?
If you want to compare departments, you need to use the `group_by()`. The `summarise()` function will then operate on the groups rather than the whole dataset:
```{r}
salaries %>% 
  group_by(department) %>% 
  summarise(overtime = sum(overtime_oncall, na.rm=TRUE)) %>% 
  arrange(desc(overtime))
```

Which job title received the most in overtime pay? 
```{r}
salaries %>% 
  group_by(job_title) %>% 
  summarise(OT = sum(overtime_oncall, na.rm=T)) %>% 
  arrange(desc(OT))
```

Explore which Senior Police Officers are getting the most overtime:
```{r}
salaries %>% filter(job_title=="Senior Police Officer") %>% 
  arrange(desc(overtime_oncall))
```

Who's getting the most overtime in terms of percentage?
To do that math we need to create a new column to store the answer for each row, and to do that we'll use the mutate() function. Using mutate is just like tacking on another column in a spreadsheet and using a formula to calculate a new value for each row.
```{r}
# test the function first:
salaries %>% 
  mutate(pct_overtime = overtime_oncall/total_comp)

# make it permanent in our environment:
salaries <- salaries %>% 
  mutate(pct_overtime = overtime_oncall/total_comp)
```

Which department has paid out the most in overtime? 
```{r}
# in raw dollars:
salaries %>% 
  group_by(department) %>% 
  summarise(total_overtime = sum(overtime_oncall, na.rm=T)) %>% 
  arrange(desc(total_overtime))

# as a percentage (per department): 
salaries %>% 
  group_by(department) %>% 
  summarise(total_overtime = sum(overtime_oncall, na.rm=T), 
            total_pay = sum(total_comp, na.rm=T)) %>% 
  mutate(overtime_pct = total_overtime/total_pay) %>% 
  arrange(desc(overtime_pct))
```

Which job_title has the highest average salary for 2021?
```{r}
salaries %>% 
  group_by(job_title) %>% 
  summarise(avg_sal = mean(salary_2021, na.rm=T), n = n()) %>% 
  arrange(desc(avg_sal))
# note that some of the high figures are job_titles with only 1 employee
```